Fastqc Analysis of Raw Reads

### 2018-10-04 notebooks/fastqc_analysis.Rmd
changed size of plots and x/y axis scales

```{r}
df_list$adapter_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    # get data ready for boxplot by gathering the columns into two columns
    gather(key,
           value, -c(ref,
                     Position,
                     group,
                     seqlen)) %>%
    # plotting
    ggplot(aes(factor(
      Position,
      levels = unique(Position),
      ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%) Adapter Content",
      color = NULL,
      title = "Adapter content"
    ) +
    scale_colour_jama() +
    scale_y_continuous(limits = c(0, 20)) +
    theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free", dir = "v")
```

## Per base sequence content
This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
```{r}
df_list$per_base_sequence_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, Base, group, seqlen)) %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%)",
      color = NULL,
      title = "Per base sequence content"
    ) +
    theme_classic() +
    scale_color_jama() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free", dir = "v")
```

@@

```{r}
df_list$sequence_duplication_levels %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, `Duplication Level`, group, seqlen)) %>%
    ggplot(aes(
      factor(
        `Duplication Level`,
        levels = unique(`Duplication Level`),
        ordered = TRUE
      ),
      value,
      fill = key
    )) +
    geom_boxplot(outlier.size = 0.5) +
    scale_fill_manual(values = c("#ef8a62",
                                 "#67a9cf")) +
    theme_classic() +
    labs(x = "Duplication Level",
         y = "Percent (%) of Sequences",
         fill = NULL,
         title = "Sequence duplication levels") +
    theme(legend.position = "bottom",
          axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.4
          )) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free")
```

@@

```{r}
df_list$per_sequence_quality_scores %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(Quality), Count, fill = factor(Quality))) +
    geom_boxplot(outlier.size = 0.5) +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Quality",
         y = "Number of reads",
         title = "Per sequence quality scores") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 10)) +
    # creates separate windows for each group
    facet_wrap( ~ group)
```

## Per sequence GC-content
This plot presents the % GC-content for all reads in each group.
```{r}
df_list$per_sequence_gc_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(`GC Content`), Count)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(x = "GC content (%)",
         y = "Number of reads",
         title = "Per sequence GC content") +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = as.character(seq(
      from = 0, to = 100, by = 10
    ))) +
    theme_classic() +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free")
```

## Per base N-content
This plot presents the per base N content for all reads in each group. N's are added to the read instead of a base if the sequencer is unable to do the base call with sufficient confidence. As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
```{r}
df_list$per_base_n_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), `N-Count`)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) +
    labs(x = "Position in read",
         title = "Per base N content") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free", dir = "v")
```

## Total deduplicated percentage
This plot presents the total deduplicated percentage for all reads in each group.
```{r}
df_list$total_deduplicated_percentage %>%
    gather(key = "ref", value = `1`) %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length",
           perc = `1`) %>%
    mutate(perc = as.numeric(perc)) %>%
    ggplot(aes(factor(group), perc)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) + 
    scale_y_continuous(limits = c(0, 100)) +
    labs(x = "Group",
         y = "Total percentage of deduplicated reads",
         title = "Total deduplicated percentage") +
    theme_classic()
```

## Per base sequence quality
This plot presents the per base mean sequence quality for all reads in each group. The green line depicts the "good quality" limit of the phred-score 28. As before, the X-axis represents the position in each read.
```{r}
df_list$per_base_sequence_quality %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    mutate(Base = factor(Base,
                         levels = unique(Base),
                         ordered = TRUE)) %>%
    group_by(group) %>%
    mutate(xmax = length(unique(Base)) + 1) %>%
    ungroup() %>%
    ggplot(aes(Base,
               Mean)) +
    geom_boxplot(outlier.size = 0.4,
                 fill = "#7f7f7f") +
    geom_hline(aes(yintercept = 28),
               color = "green") +
    labs(x = "Position in read",
         y = "Sequence quality",
         title = "Per base mean sequence quality") +
    scale_y_continuous(limits = c(0, 42)) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free", dir = "v")
```

>>>>

```{r, fig.height=8, fig.width=12}
df_list$adapter_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    # get data ready for boxplot by gathering the columns into two columns
    gather(key,
           value, -c(ref,
                     Position,
                     group,
                     seqlen)) %>%
    # plotting
    ggplot(aes(factor(
      Position,
      levels = unique(Position),
      ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%) Adapter Content",
      color = NULL,
      title = "Adapter content"
    ) +
    scale_colour_jama() +
    scale_y_continuous(limits = c(0, 20)) +
    theme_classic() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free_x", dir = "v")

```

## Per base sequence content
This plot presents the per base sequence content for all reads in each group. Each base is represented by color. 
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, Base, group, seqlen)) %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), value, color = key)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(
      x = "Position in Read",
      y = "Percent (%)",
      color = NULL,
      title = "Per base sequence content"
    ) +
    theme_classic() +
    scale_color_jama() +
    theme(
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "bottom"
    ) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free_x", dir = "v")
```

@@

```{r, fig.height=8, fig.width=12}
df_list$sequence_duplication_levels %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    gather(key, value, -c(ref, `Duplication Level`, group, seqlen)) %>%
    ggplot(aes(
      factor(
        `Duplication Level`,
        levels = unique(`Duplication Level`),
        ordered = TRUE
      ),
      value,
      fill = key
    )) +
    geom_boxplot(outlier.size = 0.5) +
    scale_fill_manual(values = c("#ef8a62",
                                 "#67a9cf")) +
    theme_classic() +
    labs(x = "Duplication Level",
         y = "Percent (%) of Sequences",
         fill = NULL,
         title = "Sequence duplication levels") +
    theme(legend.position = "bottom",
          axis.text.x = element_text(
            angle = 90,
            hjust = 1,
            vjust = 0.4
          )) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free_x")
```

@@

```{r, fig.height=8, fig.width=12}
df_list$per_sequence_quality_scores %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(Quality), Count, fill = factor(Quality))) +
    geom_boxplot(outlier.size = 0.5) +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40)) +
    scale_fill_viridis(discrete = TRUE) +
    labs(x = "Quality",
         y = "Number of reads",
         title = "Per sequence quality scores") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_text(size = 10)) +
    # creates separate windows for each group
    facet_wrap( ~ group)
```

## Per sequence GC-content
This plot presents the % GC-content for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$per_sequence_gc_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(`GC Content`), Count)) +
    geom_boxplot(outlier.size = 0.5) +
    labs(x = "GC content (%)",
         y = "Number of reads",
         title = "Per sequence GC content") +
    scale_y_continuous(labels = comma) +
    scale_x_discrete(breaks = as.character(seq(
      from = 0, to = 100, by = 10
    ))) +
    theme_classic() +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free_x")
```

## Per base N-content
This plot presents the per base N content for all reads in each group. N's are added to the read instead of a base if the sequencer is unable to do the base call with sufficient confidence. As before, the X-axis represents the position in the read, where left is the start of the read and right is the end.
```{r, fig.height=8, fig.width=12}
df_list$per_base_n_content %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    ggplot(aes(factor(
      Base, levels = unique(Base), ordered = TRUE
    ), `N-Count`)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) +
    labs(x = "Position in read",
         title = "Per base N content") +
    guides(fill = FALSE) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap( ~ group, scales = "free_x", dir = "v")
```

## Total deduplicated percentage
This plot presents the total deduplicated percentage for all reads in each group.
```{r, fig.height=8, fig.width=12}
df_list$total_deduplicated_percentage %>%
    gather(key = "ref", value = `1`) %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length",
           perc = `1`) %>%
    mutate(perc = as.numeric(perc)) %>%
    ggplot(aes(factor(group), perc)) +
    geom_boxplot(fill = "#e6e6e6",
                 outlier.size = 0.5) + 
    scale_y_continuous(limits = c(0, 100)) +
    labs(x = "Group",
         y = "Total percentage of deduplicated reads",
         title = "Total deduplicated percentage") +
    theme_classic()
```

## Per base sequence quality
This plot presents the per base mean sequence quality for all reads in each group. The green line depicts the "good quality" limit of the phred-score 28. As before, the X-axis represents the position in each read.
```{r, fig.height=8, fig.width=12}
df_list$per_base_sequence_quality %>%
    # import information from another data frame from same list,
    # to get grouping information etc.
    left_join(., df_list$basic_statistics[, c("ref", "Sequence length", "group")], by = "ref") %>%
    rename(seqlen = "Sequence length") %>%
    mutate(Base = factor(Base,
                         levels = unique(Base),
                         ordered = TRUE)) %>%
    group_by(group) %>%
    mutate(xmax = length(unique(Base)) + 1) %>%
    ungroup() %>%
    ggplot(aes(Base,
               Mean)) +
    geom_boxplot(outlier.size = 0.4,
                 fill = "#7f7f7f") +
    geom_hline(aes(yintercept = 28),
               color = "green") +
    labs(x = "Position in read",
         y = "Sequence quality",
         title = "Per base mean sequence quality") +
    scale_y_continuous(limits = c(0, 42)) +
    theme_classic() +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    # creates separate windows for each group
    facet_wrap(~ group, scales = "free_x", dir = "v")
```
